systemfit: A Package to Estimate\\ Simultaneous Equation Systems in \proglang{R}

### 2006-02-17 paper/trunk/systemfit_other.tex
Revised/extended subsection about testing linear parameter restrictions (F-test, Wald-test, LR-test)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@606 591db2bd-e0a0-4d22-8c19-2192dbb7da31

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Likelihood ratio test}

Currently, the \code{systemfit} package can perform a Likelihood Ratio
test to test linear restrictions when estimating SUR equation systems.
The function returns the empirical likelihood ratio, the $p$-value and
the degrees of freedom of the test which is equivalent to the number
of restrictions.

% This test calculates the likelihood ratio value by calculating the
% estimated residual covariance matrix $\hat{S}$ by the maximum
% likelihood formula,

% \begin{equation}
% \hat{s}_{ij} = \frac{\hat{e}_i^{'} \hat{e}_j}{T} 
% \end{equation}
% where $T$ is vector of ones.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Testing Linear Restrictions}

Linear restrictions can be tested by an F test, Wald test or
likelihood-ratio (LR) test.

The F-statistic for sytems of equations is
\begin{equation}
F = \frac{
   ( R \hat{b} - q )'
   ( R ( X' ( \hat{\Sigma} \otimes I )^{-1} X )^{-1} R' )^{-1}
   ( R \hat{b} - q ) /
   j
}{
   \hat{e}' ( \Sigma \otimes I )^{-1} \hat{e} /
   ( M \cdot T - K )
}
\end{equation}
where $j$ is the number of restrictions,
$M$ is the number of equations,
$T$ is the number of observations per equation,
$K$ is the total number of estimated coefficients, and
$\hat{\Sigma}$ is the estimated residual covariance matrix
used in the estimation.
Asymptotically, $F$ has an F-distribution
with $j$ and $M \cdot T - K$ degrees of freedom
under the null hypothesis
\citep[p.\ 346]{greene03}.

The Wald-statistic for sytems of equations is
\begin{equation}
W =
   ( R \hat{b} - q )'
   ( R \widehat{Cov} [ \hat{b} ] R' )^{-1}
   ( R \hat{b} - q )
\end{equation}
Asymptotically, $W$ has a $\chi^2$
distribution with $j$ degrees of freedom
under the null hypothesis
\citep[p.\ 347]{greene03}.

The LR-statistic for sytems of equations is
\begin{equation}
LR = T \cdot \left(
   log \left| \hat{ \hat{ \Sigma } }_r \right|
   - log \left| \hat{ \hat{ \Sigma } }_u \right|
   \right)
\end{equation}
where $T$ is the number of observations per equation, and 
$\hat{\hat{\Sigma}}_r$ and $\hat{\hat{\Sigma}}_u$ are
the residual covariance matrices calculated by formula (\ref{eq:rcov0})
of the restricted and unrestricted estimation, respectively.
Asymptotically, $LR$ has a $\chi^2$
distribution with $j$ degrees of freedom
under the null hypothesis
\citep[p.\ 349]{greene03}.
